---
title: "CS-767 Assignment 2"
author: "Benzon Carlitos Salazar"
date: "2024-02-18"
output: pdf_document
---

```{r setup, include=FALSE}
library(magrittr)
```

## Read Data

```{r read_data}
movie_data_df <-
  readr::read_csv(here::here("data", "movie_data.csv"), show_col_types = FALSE)
summary(movie_data_df)

rt_movie_data_df <-
  readr::read_csv(here::here("data", "rt_movie_data.csv"), show_col_types = FALSE)
summary(rt_movie_data_df)
```

## Schema definition

In both tables, each row represents a movie, with columns representing different attributes such as ID, title, year, 
rating, and rank.

### `movie_data.csv`

```{r movie_data_schema}
movie_data_df %>%
  dplyr::slice_head(n = 10) %>%
  knitr::kable()
```

- ID: UUID (Universally Unique Identifier)
- Title: String (Movie title)
- Year: Integer (Year of release)
- Rating: Float (Movie rating)
- Rank: Integer (Ranking of the movie)

### `rt_movie_data.csv`

```{r rt_movie_data_schema}
rt_movie_data_df %>%
  dplyr::slice_head(n = 10) %>%
  knitr::kable()
```

- ID: UUID (Universally Unique Identifier)
- Title: String (Movie title)
- Year: Integer (Year of release)
- Rating: Integer (Movie rating)
- Rank: Integer (Ranking of the movie)

## Missing Values

### `movie_data.csv` schema

```{r missing_values_for_movie_data}
missin_values_movie_data <-
  movie_data_df %>%
  dplyr::summarise_all(list(~ sum(is.na(.))))

missin_values_movie_data %>%
  knitr::kable()
```

### `rt_movie_data.csv` schema

```{r missing_values_for_rt_movie_data}
missin_values_rt_movie_data <-
  rt_movie_data_df %>%
  dplyr::summarise_all(list(~ sum(is.na(.))))

missin_values_rt_movie_data %>%
  knitr::kable()
```

There are no missing data in both tables. Each column in the tables contains values for all rows, with no instances 
where a value is missing (i.e., no NA or NULL values). This indicates that the data is complete and there are no missing 
values to be addressed in the analysis.

If there happens to be missing values then,

- for the `ID` column, theoretically, this column does not allow any NULL values because this is an identifier. If these 
were missing, this tells us that there are no corresponding movies, in which case we would just remove these rows.
- for the `Title` column, because we have no way of determining what movie the attributes would correspond to, we can just 
use a default value of `Unknown Movie` or something similar.
- for the `Year` column, since these are numeric values, we can use a default `1601` value to indicate any missing values. In this 
case, since we know that no good movies existed in year 1601, it is a good indicator of missing values.
- for the `Rating` column, we may use the average for that column to fill in missing values. Since the table is ordered 
by ranking, we may get away with finding a somewhat correct rating based on the rows position.
- for the `Rank` column, since the rows are already ordered, we can simply use the rows index to fill in any missing 
values in the `Rank` column.

## Column classification

- ID: UUID (Universally Unique Identifier)
- Title: String (Movie title)
- Year: Integer (Year of release)
- Rating: Integer (Movie rating)
- Rank: Integer (Ranking of the movie)

## Character Column Attribution Report for `Title` column in `movie_data.csv`

```{r helper_function}
# Helper function to calculate average, min, and max length of textual values
textual_stats <- function(x) {
  if(is.character(x)) {
    avg_length <- mean(nchar(x), na.rm = TRUE)
    min_length <- min(nchar(x), na.rm = TRUE)
    max_length <- max(nchar(x), na.rm = TRUE)
    return(data.frame(avg_length = avg_length, min_length = min_length, max_length = max_length))
  } else {
    return(NULL)
  }
}
```


```{r movie_data_col_title}
movie_data_summary <-
  movie_data_df %>%
  dplyr::summarise(dplyr::across(Title, textual_stats))
```

- Average length of characters in `Title` column: `r movie_data_summary$Title$avg_length` characters.
- Minimum length of characters in `Title` column: `r movie_data_summary$Title$min_length` character.
- Maximum length of characters in `Title` column: `r movie_data_summary$Title$max_length` characters.

## Character Column Attribution Report for `Title` column in `rt_`movie_data.csv`

```{r rt_movie_data_col_title}
rt_movie_data_summary <-
  rt_movie_data_df %>%
  dplyr::summarise(dplyr::across(Title, textual_stats))
```

- Average length of characters in `Title` column: `r rt_movie_data_summary$Title$avg_length` characters.
- Minimum length of characters in `Title` column: `r rt_movie_data_summary$Title$min_length` character.
- Maximum length of characters in `Title` column: `r rt_movie_data_summary$Title$max_length` characters.

## Outliers

### Histogram for the Rating column in `movie_data.csv`

```{r outliers_movie_data}
ggplot2::ggplot(movie_data_df, ggplot2::aes(x = Rating)) +
  ggplot2::geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  ggplot2::labs(title = "Distribution of Ratings", x = "Rating", y = "Frequency")
```

### Histogram for the length of the Title column in `movie_data.csv`

```{r outliers_title_movie_data}
movie_data_df$Title_length <- nchar(movie_data_df$Title)

ggplot2::ggplot(movie_data_df, ggplot2::aes(x = Title_length)) +
  ggplot2::geom_histogram(binwidth = 5, fill = "lightgreen", color = "black") +
  ggplot2::labs(title = "Distribution of Title Lengths", x = "Length", y = "Frequency")
```

### Histogram for the Rating column in `rt_movie_data.csv`

```{r outliers_rt_movie_data}
ggplot2::ggplot(rt_movie_data_df, ggplot2::aes(x = Rating)) +
  ggplot2::geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  ggplot2::labs(title = "Distribution of Ratings", x = "Rating", y = "Frequency")
```

### Histogram for the length of the Title column in `movie_data.csv`

```{r outliers_title_rt_movie_data}
rt_movie_data_df$Title_length <- nchar(rt_movie_data_df$Title)

ggplot2::ggplot(rt_movie_data_df, ggplot2::aes(x = Title_length)) +
  ggplot2::geom_histogram(binwidth = 5, fill = "lightgreen", color = "black") +
  ggplot2::labs(title = "Distribution of Title Lengths", x = "Length", y = "Frequency")
```

We don't really see much outliers, except for the length of characters in titles. The ratings seem to be correctly 
formatted and are within the range. The only difference is the fact that ratings in `movie_data.csv` are in the range of 
1 - 10, while the ratings in the `rt_movie_data.csv` are in the range of 0 - 100.

We do see ratings around 70 - 75 in the `rt_movie_data.csv`, and quite a few of 100s in the `rt_movie_data.csv`. What we 
can do to match the range of ratings for both dataframes is to divide each rating in the `rt_movie_data.csv` by 10 to 
standardize it to be between 0 - 10. Then, we can re-run the histogram and see if it worked.

```{r standardize_rt}
rt_movie_data_df <-
  rt_movie_data_df %>%
  dplyr::mutate(Rating = Rating / 10)

ggplot2::ggplot(rt_movie_data_df, ggplot2::aes(x = Rating)) +
  ggplot2::geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  ggplot2::labs(title = "Distribution of Standardized Ratings", x = "Rating", y = "Frequency")
```

After standardization, we now see that the majority of ratings are in the 9.5 range.

## Formatting

All the columns do follow the same formatting. For example, all of the `Year` values are properly formatted as `YYY`. 
The `Rank` values are all integers, and so are the `Rating` values.

## Additional Data Quality Problems

Upon initial review, both datasets appear to be free of inherent data quality issues. The only discrepancy noted is the 
difference in ratings between the two tables, indicating a potential inconsistency in the movies listed. This suggests 
that one table may include movies not present in the other, highlighting a need for further investigation to ensure data 
completeness and consistency across the datasets.

## Packages used

The following R packages are used in this report:

```{r packages_used, echo=FALSE}
packages <- renv::dependencies()$Package %>% as.data.frame()
packages %>% dplyr::distinct() %>% as.list()
```

```{r clean, include=FALSE, echo=FALSE}
rm(list = ls(all.names = TRUE)) # clear all objects including hidden objects
invisible(gc()) # free up memory
```

